---
title: "Rockies Final Project"
author: "Cassy Miller"
date: "5/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r knitr-options, include = FALSE}
knitr::opts_chunk$set(fig.align="center",
                      warning = FALSE,
                      message = FALSE,
                      comment = NA)
```

Toronto Blue Jays at Colorado Rockies, May 31, 2019
```{r, include = FALSE}
library(tidyverse)
library(ggcorrplot)
library(baseballr)
library(lubridate)
library(mgcv)
library(gganimate)
library(writexl)
library(rvest)
theme_set(theme_classic())

```

Looking at the Rockies Batters For the 19' Season
```{r,include=FALSE}
R_Tapia_id <- playerid_lookup(last_name = "Tapia", first_name = "Raimel")%>%
  dplyr::pull(.,mlbam_id)
R_Tapia <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = R_Tapia_id)
T_Story_id <- playerid_lookup(last_name = "Story", first_name = "Trevor")%>%
  dplyr::pull(.,mlbam_id)
T_Story <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = T_Story_id)
D_Dahl_id <- playerid_lookup(last_name = "Dahl", first_name = "David")%>%
  dplyr::pull(.,mlbam_id)
D_Dahl <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = D_Dahl_id)
B_Rodgers_id <- playerid_lookup(last_name = "Rodgers", first_name = "Brendan")%>%
  dplyr::pull(.,mlbam_id)
B_Rodgers <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = B_Rodgers_id)
D_Murphy_id <- playerid_lookup(last_name = "Murphy", first_name = "Daniel")
D_Murphy <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = "502517")
R_Mcmahon_id <- playerid_lookup(last_name = "Mcmahon", first_name = "Ryan")
R_Mcmahon <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = "641857")
I_Desmond_id <- playerid_lookup(last_name = "Desmond", first_name = "Ian")%>%
  dplyr::pull(.,mlbam_id)
I_Desmond <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = I_Desmond_id)
T_Wolters_id <- playerid_lookup(last_name = "Wolters", first_name = "Tony")%>%
  dplyr::pull(.,mlbam_id)
T_Wolters <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = T_Wolters_id)
J_Tinoco_id <- playerid_lookup(last_name = "Tinoco", first_name = "Jesus")%>%
  dplyr::pull(.,mlbam_id)
J_Tinoco <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = J_Tinoco_id)
N_Arenado_id <- playerid_lookup(last_name = "Arenado", first_name = "Nolan")%>%
  dplyr::pull(.,mlbam_id)
N_Arenado <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = N_Arenado_id)
M_Reynolds_id <- playerid_lookup(last_name = "Reynolds", first_name = "Mark")
M_Reynolds <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = "448602")
C_Iannetta_id <- playerid_lookup(last_name = "Iannetta", first_name = "Chris")%>%
  dplyr::pull(.,mlbam_id)
C_Iannetta <- scrape_statcast_savant(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       playerid = C_Iannetta_id)

```
```{r, include=FALSE}
write_xlsx(R_Tapia,"C:\\Users\\Cassy\\Desktop\\1.xlsx")
write_xlsx(T_Story,"C:\\Users\\Cassy\\Desktop\\2.xlsx")
write_xlsx(D_Dahl,"C:\\Users\\Cassy\\Desktop\\3.xlsx")
write_xlsx(B_Rodgers,"C:\\Users\\Cassy\\Desktop\\4.xlsx")
write_xlsx(D_Murphy,"C:\\Users\\Cassy\\Desktop\\5.xlsx")
write_xlsx(R_Mcmahon,"C:\\Users\\Cassy\\Desktop\\6.xlsx")
write_xlsx(I_Desmond,"C:\\Users\\Cassy\\Desktop\\7.xlsx")
write_xlsx(T_Wolters,"C:\\Users\\Cassy\\Desktop\\8.xlsx")
write_xlsx(J_Tinoco,"C:\\Users\\Cassy\\Desktop\\9.xlsx")
write_xlsx(N_Arenado,"C:\\Users\\Cassy\\Desktop\\10.xlsx")
write_xlsx(M_Reynolds,"C:\\Users\\Cassy\\Desktop\\11.xlsx")
write_xlsx(C_Iannetta,"C:\\Users\\Cassy\\Desktop\\12.xlsx")
```


```{r}
massive_df <- read.csv("~/DU/Junior Year/Spring 2020/Sports Amalytics/Rockies Project/massive_data.csv")

```
```{r}
plots <- massive_df %>% 
  mutate(., hc_yy = (-1*hc_y)) %>%
  mutate(., out_v = if_else(events == c("single", "home_run", "double", "triple"), "on_base", "out"))
plots31 <- plots %>%
  filter(., game_date == "2019-05-31")

```
```{r}
events <- tbl_df(unique(massive_df$events)) 
events
type <- tbl_df(unique(massive_df$type))
type
```
Looking at Player's 2019 spraycharts
```{r}
p1 <- ggspraychart(data= plots, point_alpha =.5, x_value = "hc_x", y_value = "hc_yy")+ 
  
  theme(legend.position = "none",plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F")) +
  transition_manual(player_name) +labs(title= "{current_frame}'s 2019 hits") 

animate(p1, fps = 5)


```
Looking at Player's spraycharts for May 31, 2019
```{r}
p2 <- ggspraychart(data= plots31, point_alpha =.5, x_value = "hc_x", y_value = "hc_yy") +
  theme(legend.position = "none",plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F")) +
  transition_manual(player_name) +labs(title= "{current_frame}'s hits for May 31, 2019 V TOR") 

animate(p2, fps = 5)

```

```{r, include=FALSE}
top_zone <- 3.5
bot_zone <- 1.6
left_zone <- -0.75
right_zone <- 0.75
strike_zone_df <- data.frame(
  x = c(left_zone, left_zone, right_zone, right_zone, left_zone),
  y = c(bot_zone, top_zone, top_zone, bot_zone, bot_zone)
)
```
## Batter's Box for 2019 Season

```{r}
swing_set <- plots %>% 
  mutate(swing = if_else(description%in%
                        c("ball","blocked_ball","automatic_ball", "called_strike", "hit_by_pitch"), 0 , 1))

p3 <- ggplot(swing_set,  
             aes(x = plate_x, y = plate_z, col = as.factor(swing))) + geom_point(alpha = 0.2)  + scale_color_brewer("swing", palette = "Set1") +
  geom_path(data = strike_zone_df, aes(x, y), lwd = 1, color = "black") + 
  xlim(-3.5, 3.5) + 
  coord_fixed() + transition_manual(player_name) +labs(title =  "{current_frame}'s batter box for 2019 season", x= "horizontal location", y = "vertical location") +theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F"))
animate(p3, fps =2)
```


```{r}
swing_set31 <- swing_set %>%
  filter(.,game_date=="2019-05-31")
p4 <- ggplot(swing_set31,  
             aes(x = plate_x, y = plate_z, col = as.factor(swing))) + geom_point(alpha = 0.15)  + scale_color_brewer("swing", palette = "Set1") +
  geom_path(data = strike_zone_df, aes(x, y), color = "black") + 
  xlim(-3.5, 3.5) + 
  coord_fixed() + transition_manual(player_name) +labs(title =  "{current_frame}'s batter box for May 31, 2019 v TOR", x= "horizontal location", y = "vertical location")  +theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F"))
animate(p4, fps =2)

```
```{r}

p5 <- ggplot(data =swing_set31, 
             aes(x = plate_x, y = plate_z, col = as.factor(swing), text = paste('Pitch:',swing_set31$pitch_type, '\n', 'Outcome:',events,",",swing_set31$description,'\n'))) +  geom_point(alpha = 0.15) +
  facet_wrap(~player_name, ncol=4)+ scale_color_brewer("swing", palette = "Set1")  
plotly::ggplotly(p5, tooltip="text") 
```
```{r}
swing_set31_1<- swing_set31 %>%
  filter(swing=="1")
p5 <- ggplot(data = swing_set31_1,
             aes(x = plate_x, y= plate_z, col=player_name, text = paste(player_name, '\n', 'Pitch: ', pitch_type, "\n", "Outcome: ", events, ",", description))) + scale_color_brewer("player_name",palette="Set3")+
  geom_point(alpha=.5) + labs(title = "Swings during May 31, 2019 v TOR", x="Horizontal Location", y= "Vertical location") +theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F")) + coord_fixed()

plotly::ggplotly(p5, tooltip = "text")
```


+ geom_path(data = strike_zone_df, aes(x, y), lwd = 1, color = "black") + 
  xlim(-3.5, 3.5) 
  coord_fixed()
  
Faced Pitches by player
```{r}
faced_pit31 <- massive_df %>% 
  filter(!pitch_type == "null") %>% filter( game_date== "2019-05-31") 

p6 <- ggplot(data= faced_pit31, aes(x=pitch_type))
p6 + geom_bar(aes(fill=player_name)) +labs(title = "Pitches the Rockies Batters Faced May 31, 2019 v TOR", x= "Pitch Type", y= "Count") + scale_fill_brewer(palette = "Set3") +theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F")) 
  

```



Faced pitches by pitcher

```{r}
p7 <- ggplot(data= faced_pit31, aes(x=pitch_type))
p7 + geom_bar(fill="#33006F") +labs(title = "Pitches the Rockies Batters Faced May 31, 2019 v TOR by pitcher", x= "Pitch Type", y= "Count")  +theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F")) +facet_grid("pitcher...8")
```
Comparing hit distance and launch angle
```{r}
hits <- plots %>%
  filter(., type == "X") %>%
  mutate(may31 = if_else(game_date=="2019-05-31", "May 31, 2019", "'19 Season"))

p8 <- ggplot(data = hits,
            aes(x = launch_angle, y = hit_distance_sc, 
                text=paste('Player: ', player_name,'\n', 'Pitch Type: ', pitch_type, '\n', events))) +
  geom_point(color = "#33006F",alpha = .15) +
  labs(title = "Hit Distance v Launch Angle", description="Comparing 2019 season to May 31, 2019 v TOR", x = "Launch Angle", y= "Hit Distance") +facet_wrap("may31") +theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F"))

plotly::ggplotly(p8, tooltip = "text")
```
Comparing Launch Speed and Hit Distance
```{r}
p9 <- ggplot(data = hits,
            aes(x = launch_speed, y = hit_distance_sc, 
                text=paste('Player: ', player_name,'\n', 'Pitch Type: ', pitch_type, '\n', events))) +
  geom_point(color = "#33006F",alpha = .15) +
  labs(title = "Hit Distance v Launch Speed", description="Comparing 2019 season to May 31, 2019 v TOR", x = "Launch Speed", y= "Hit Distance") +facet_wrap("may31") +theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F"))

plotly::ggplotly(p9, tooltip = "text")
```


+ transition_manual(player_name)
gganimate::animate(p8, fps=3)

Player's Stat Charts 
```{r}
url_base <- "https://www.baseball-reference.com/boxes/COL/COL201905310.shtml"


#statcharts <- xml2::read_html(url) %>%
 # rvest::html_nodes(xpath= '//*[(@id = "ColoradoRockiesbatting")]') %>%
  #rvest::html_table()

```
  links <- xml2::read_html(url) %>%
  rvest::html_nodes(xpath = '//*[(@id = "roster")]//a') %>%
  rvest::html_attr('href')
links

Looking at the batter's box scores
```{r}
url_base <- "http://gd2.mlb.com/components/game/mlb/"
 url <- paste0(url_base, 
   "year_2019/month_05/day_31/gid_2019_05_31_milmlb_nynmlb_1/boxscore.xml")
 batter_boxscore(url)

```


```{r}
df1 <-get_game_pks_mlb("2019-05-31",level_ids = c(1)) %>%
  filter(teams.home.team.name == "Colorado Rockies")
df1


```
```{r}
order_COL <- get_batting_orders(game_pk = df1$game_pk, type = "starting") %>%
  filter(team=="home") %>%
  select("id", "fullName", "abbreviation", "batting_order")
order_TOR <- get_batting_orders(game_pk = df1$game_pk, type = "starting") %>%
  filter(team=="away") %>%
  select("id", "fullName", "abbreviation", "batting_order")
```
Colorado's Batters
```{r}
print(order_COL)
```

Toronto's Batters
```{r}
print(order_TOR)
```
Game Info
```{r}
game31 <- get_game_info_mlb(game_pk = df1$game_pk)
print(game31)
```
pbp
```{r}
pbp<-get_pbp_mlb(game_pk = df1$game_pk)
pbp
```

## Batters Box score
```{r}
url_base <- "http://gd2.mlb.com/components/game/mlb/"
url <- paste0(url_base, 
  "year_2019/month_05/day_31/gid_2019_05_31_tormlb_colmlb_1/boxscore.xml")
batbox <- batter_boxscore(url)
batbox
```
## Charting The Data
```{r}
onb <- batbox %>%
  select(name_display_first_last, obp, rbi, avg ) %>% mutate(t = if_else(batbox$team == "21", "TOR", "COL"))
onb

onbplot <- ggplot(onb,
                  aes(x= obp, y= avg, col=t, text = 
                        paste('Player:', name_display_first_last, '\nOn Base Percent:', obp, '\nBatting Average:', avg, '\nRuns Batted In:', rbi ))) +  
  geom_point() +
  scale_color_brewer(palette = "Set1") + labs(title = "Comparing OBP to AVG", x= "On Base Percentage (obp)", y= "Batting Average (avg)")  + theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F"))

plotly::ggplotly(onbplot, tooltip = "text")
```



Pitchers
```{r}
pitbox <- pitcher_boxscore(url) %>% select(!"game_score")
pitbox
```

## Analyzing COL Pitching 

```{r}
Marquez_id <- playerid_lookup(last_name = "Marquez", first_name = "German")%>%
  dplyr::pull(.,mlbam_id)
Marquez <- scrape_statcast_savant_pitcher(start_date = ymd("2019-03-28"), 
                                       end_date = ymd("2019-09-29"), 
                                       pitcherid = Marquez_id)

```
## Marquez's pitches by Type
```{r}
pitches <- ggplot(data= Marquez, 
                  aes(x= pitch_name)) +
  geom_bar(fill="#33006F") + labs(title="German Marquez's pitches for 2019 season", x= "Pitch Type", y= "Count") +
  theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F"))
pitches
```

```{r}
Marquez31 <- Marquez %>%
  filter(.,game_date == "2019-05-31")
pitches2 <- ggplot(data= Marquez31, 
                  aes(x= pitch_name)) +
  geom_bar(fill="#33006F") + labs(title="German Marquez's pitches game May 31,2019 v TOR", x= "Pitch Type", y= "Count") +
  theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F"))
pitches2

```

## Release Speed by Pitch Type

```{r}
rs_pitch <- Marquez %>%
filter(type == "X") %>%
ggplot(aes(x = pitch_type, y = release_speed, fill = pitch_name)) + 
  geom_violin(alpha = 0.5, color = "black") +
scale_fill_brewer("type", palette = "BuPu") +
labs(title = "German Marquez's Release Speed by Pitch Type in 2019",
x = "Pitch Type",
y = "Release Speed")  +
  theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F")) + 
  theme(legend.position = "none")+ 
coord_flip()
rs_pitch

```

```{r}
rs_pitch2 <- Marquez31 %>%
filter(type == "X") %>%
ggplot(aes(x = pitch_type, y = release_speed, fill = pitch_name)) + 
  geom_violin(alpha = 0.5, color = "black") +
scale_fill_brewer("type", palette = "BuPu") +
labs(title = "German Marquez's Release Speed by Pitch Type \nMay 31, 2019 v TOR",
x = "Pitch Type",
y = "Release Speed")  + 
  theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F"))+ 
  theme(legend.position = "none")+ 
coord_flip()
rs_pitch2
```

## Pitching Accuracy

```{r}
acc <- ggplot(Marquez,  
             aes(x = plate_x, y = plate_z, col = type )) + geom_point(alpha = 0.4)  + scale_color_brewer("type", palette = "Set1")  + 
  
  geom_path(data = strike_zone_df, aes(x, y), lwd = 1, color = "black") + 
  xlim(-3.5, 3.5) +
  coord_fixed() + transition_manual(pitch_name) +labs(title =  " Marquez's {current_frame} pitches in 2019", x= "horizontal location", y = "vertical location") +theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F"))
animate(acc, fps =5)
```

```{r}
acc31 <- ggplot(Marquez31,  
             aes(x = plate_x, y = plate_z, col = type )) + geom_point(alpha = 0.4)  + scale_color_brewer("type", palette = "Set1")  + 
  
  geom_path(data = strike_zone_df, aes(x, y), lwd = 1, color = "black") + 
  xlim(-3.5, 3.5) +
  coord_fixed() + transition_manual(pitch_name) +labs(title =  " Marquez's {current_frame} pitches on \nMay 31, 2019 v TOR", x= "horizontal location", y = "vertical location") +theme(plot.title = element_text(hjust =.5, size=15, face = "bold", color = "#33006F"))
animate(acc31, fps =5)
```

## Speed of Pitches

```{r}
url_base <- "http://gd2.mlb.com/components/game/mlb/"
url <- paste0(url_base, 
  "year_2019/month_05/day_31/gid_2019_05_31_tormlb_colmlb_1/boxscore.xml")
pitbox <- pitcher_boxscore(url) %>% select(.,!one_of("game_score"))
pitbox
```




























##########################
```{r Tapia31}
R_Tapia_bat31 <- R_Tapia_bat %>% filter(., game_date=="2019-05-31")
p1 <- ggplot(R_Tapia_bat31, 
             aes(x = plate_x, y = plate_z, col = as.factor(swing)))
p1 + geom_point() +
  scale_color_brewer("swing", palette = "Set1") +
  geom_path(data = strike_zone_df, aes(x, y), lwd = 1, color = "black") +
  labs(title = "R. Tapia's Hits for game May 31, 2019 vs. TOR
       ",
       x = "horizontal location (ft)",
       y = "vertical location (ft)") +
  xlim(-3.5, 3.5) +
  coord_fixed()


```

```{r}
R_Tapia_bat31 <- R_Tapia_bat31 %>% 
  mutate(., hc_yy = (-1*hc_y))
ggspraychart(data= R_Tapia_bat31, point_alpha =.5, x_value = "hc_x", y_value = "hc_yy", fill_value = "events", fill_palette = c("field_out"="red","force_out"="red","single"="#33006F", "double"="#33006F", "triple"="#33006F", "home_run"="#33006F") ) +
  ggtitle("\nR. Tapia Hits May 31, 2019 vs TOR") + theme(legend.position = "none",plot.title = element_text(hjust =.5, size=15, face = "bold"))
```
```{r}


```
Tapia's Hits
```{r}
R_Tapia_bat <- R_Tapia_bat %>% 
  mutate(., hc_yy = (-1*hc_y))
ggspraychart(data= R_Tapia_bat, point_alpha =.5, x_value = "hc_x", y_value = "hc_yy") +
  ggtitle("\nR. Tapia Hits 2019") + theme(legend.position = "none",plot.title = element_text(hjust =.5, size=15, face = "bold"))
```

```{r}
R_Tapia_bat <- R_Tapia_bat %>% 
  mutate(., hc_yy = (-1*hc_y))
ggspraychart(data= R_Tapia_bat, point_alpha =.5, x_value = "hc_x", y_value = "hc_yy", fill_value = "events", fill_palette = c("field_out"="red","force_out"="red","single"="#33006F", "double"="#33006F", "triple"="#33006F", "home_run"="#33006F", "strikeout" ="red", "field_error"="red", "walk"="white", "strikeout_double_play" = "green", "grounded_into_double_play"="#33006f", "null"="yellow") ) +
  ggtitle("\nR. Tapia Hits 2019") + theme(legend.position = "none",plot.title = element_text(hjust =.5, size=15, face = "bold"))
```

```{r}

p2 <- ggplot(data =R_Tapia_bat31, aes(x= pitch_type))

p2 + geom_bar() +labs(title = "Pitches R. Tapia say May 31, 2019") + geom_bar(data=R_Tapia_bat, aes(x=pitch_type))
```
